Using Model Checking to Analyze the System Behavior of the LHC Production Grid

### 2013-02-11 ElsevierFGCS_singleColumn.tex
Related work...

The development of DIRAC started in 2002 as a system for production of simulation
data that would serve to verify physics theory, aspects of the LHCb detector design, as
well as to optimize physics algorithms. It gradually evolved into a complete community
grid solution for data and job management, based on a general-purpose framework
that can be reused by other communities besides LHCb. Today, it covers all major
LHCb tasks starting with the raw data transfer from the experiment's detector to
the grid storage, several steps of data processing, up to the final user
analysis. 
It provides a concurrent use of over 60K CPUs and 10M file replicas distributed over
tens of grid sites, provided by the WLCG computing grid \cite{Bird:2011zz}. The community of users has grown to several hundreds, loading the 
LHC grid resources with up to 30K simultaneously running jobs (Fig.~\ref{fig:LHCb-resource-usage}, top) and over 100000 jobs executed per day, during peak processing periods.
About 10PB of data is stored and visible in the LHCb file catalogs, with daily data transfers among storage units reaching 3Gbps (Fig.~\ref{fig:LHCb-resource-usage}, bottom).
\begin{figure*}[tp!]
  \centering
  {\label{fig:Job-throughput}\includegraphics[width=0.7\linewidth]{./graphics/Figure1_1.png}}    
  \vfill
  \vspace{14pt}
{\label{fig:Data-transfers}\includegraphics[width =0.7\linewidth]{./graphics/Figure1_2.png}}
  \caption{LHCb resource usage}
  \label{fig:LHCb-resource-usage}
\end{figure*}
Python was chosen as the implementation language, since it enables
rapid prototyping and development of new features. DIRAC follows the Service
Oriented Architecture (SOA) paradigm, accompanied by a network of lightweight
distributed agents which animate the system. Its main components are depicted in
Fig.~\ref{fig:DIRAC-Arch}.

@@

A simplified view of the system is shown in  Fig.~\ref{fig:DIRAC-SMS}. The workflow is initiated
with the \textit{Job Scheduling Agent} detecting that a job is assigned to process files
only available on tape storage. It sends a request for staging (i.e., creating a
cached replica) to the \textit{Storage Manager Handler} service with the
list of files and a callback method to be invoked when the request has been
processed. The Storage Management DB is immediately populated with records which
are processed by a sequence of agents in an organized fashion. The relevant
tables in the SMS DB are the \textit{Tasks} and \textit{CacheReplicas}, whose
entities maintain a state observed and updated by these agents. \textit{Tasks} maintain
general information about every job requesting a service from the SMS. The
details about every file (i.e., the Storage Element where it resides, the
size, checksum, number of tasks that requested it), are kept in the \textit{CacheReplicas}
table. Other auxiliary tables maintain the relationship between these entities.
\begin{figure}[tp]
\begin{center}
 \includegraphics[width=0.5\linewidth,keepaspectratio=true]{./graphics/Figure4.png}
\end{center}
\caption{DIRAC Storage Management System}
\label{fig:DIRAC-SMS}
\end{figure}
The processing begins with the \textit{Request Preparation Agent}. It selects all the
``New'' replica entries, checks whether they are registered in a Logical
File Catalog (LFC), and retrieves their metadata. In case of problematic catalog
entries, it can update the state of the \textit{CacheReplicas} and the related \textit{Tasks}
entries to ``Failed''. Non-problematic files are updated to a
``Waiting'' state. The \textit{Stage Request Agent} is responsible for placing the actual staging
requests for all ``Waiting'' entries, via dedicated storage middleware that
communicates with the tape backends. These requests are grouped by Storage Element (SE) prior to
submission, and carry information about the requested (pin) lifetime of the
replicas to be cached. If certain pathologies are discovered (i.e., lost,
unavailable, or zero-sized files on tape), it can update the corresponding
entries to ``Failed'' in a similar manner. Otherwise, they are promoted to
``StageSubmitted''. The agent responsible for monitoring the status of submitted
requests is the \textit{Stage Monitor Agent}. It achieves this by interrogating the
storage middleware to see if the ``StageSubmitted'' files are successfully
replicated on disk cache. In case of success, the \textit{CacheReplicas} and their
corresponding \textit{Tasks} entries are updated to ``Staged''. Various circumstances of
tape or middleware misbehavior can also fail the staging requests. The last one
in the chain is the \textit{Request Finalization Agent}. The \textit{Tasks} which are in their
final states (``Staged'' or ``Failed'') are cleared from the database, and callbacks
are performed to the WMS, which effectively wakes up the corresponding jobs. If
there are no more associated \textit{Tasks} for particular replicas, the respective
\textit{CacheReplicas} entries are also removed.

@@

>>>>

The development of DIRAC started in 2002 as a system for production of simulation
data that would serve to verify physics theory, aspects of the LHCb detector design, as
well as to optimize physics algorithms. It gradually evolved into a complete community
grid solution for data and job management, based on a general-purpose framework
that can be reused by other communities besides LHCb. Today, it covers all major
LHCb tasks starting with the raw data transfer from the experiment's detector to
the grid storage, several steps of data processing, up to the final user
analysis. 
It provides a concurrent use of over 60K CPUs and 10M file replicas distributed over
tens of grid sites, provided by the WLCG computing grid \cite{Bird:2011zz}. The community of users has grown to several hundreds, loading the 
LHC grid resources with up to 30K simultaneously running jobs (Fig.~\ref{fig:LHCb-resource-usage}, top) and over 100K jobs executed per day, during peak processing periods.
About 10PB of data is stored and visible in the LHCb file catalogs, with daily data transfers between storage units reaching 3Gbps (Fig.~\ref{fig:LHCb-resource-usage}, bottom).
\begin{figure*}[tp!]
  \centering
  {\label{fig:Job-throughput}\includegraphics[width=0.7\linewidth]{./graphics/Figure1_1.png}}    
  \vfill
  \vspace{14pt}
{\label{fig:Data-transfers}\includegraphics[width =0.7\linewidth]{./graphics/Figure1_2.png}}
  \caption{LHCb resource usage}
  \label{fig:LHCb-resource-usage}
\end{figure*}
Python was chosen as the implementation language, since it enables
rapid prototyping and development of new features. DIRAC follows the Service
Oriented Architecture (SOA) paradigm, accompanied by a network of lightweight
distributed agents which animate the system. Its main components are depicted in
Fig.~\ref{fig:DIRAC-Arch}.

@@

A simplified view of the system is shown in  Fig.~\ref{fig:DIRAC-SMS}. The workflow is initiated
with the \textit{Job Scheduling Agent} detecting that a job is assigned to process files
only available on tape storage. It sends a request for staging (i.e., creating a
cached replica) to the \textit{Storage Manager Handler} service with the
list of files and a callback method to be invoked when the request has been
processed. The Storage Management DB is immediately populated with records which
are processed by a sequence of agents in an organized fashion. The relevant
tables in the SMS DB are the \textit{Tasks} and \textit{CacheReplicas}, whose
entities maintain a state observed and updated by these agents. \textit{Tasks} maintain
general information about every job requesting a service from the SMS. The
details about every file (i.e., the Storage Element where it resides, the
size, checksum, number of tasks that requested it), are kept in the \textit{CacheReplicas}
table. Other auxiliary tables maintain the relationship between these entities.

\begin{figure}
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=0.98\linewidth]{./graphics/Figure4.png}
  \caption{DIRAC Storage Management System}
  \label{fig:DIRAC-SMS}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=0.95\linewidth]{./graphics/SMS_CacheReplicasSM.png}
  \caption{CacheReplicas state machine}
  \label{fig:CacheReplicas-SM}
\end{minipage}
\end{figure}

The processing begins with the \textit{Request Preparation Agent}. It selects all the
``New'' replica entries, checks whether they are registered in a Logical
File Catalog (LFC), and retrieves their metadata. In case of problematic catalog
entries, it can update the state of the \textit{CacheReplicas} and the related \textit{Tasks}
entries to ``Failed''. Non-problematic files are updated to a
``Waiting'' state. The \textit{Stage Request Agent} is responsible for placing the actual staging
requests for all ``Waiting'' entries, via dedicated storage middleware that
communicates with the tape backends. These requests are grouped by Storage Element (SE) prior to
submission, and carry information about the requested (pin) lifetime of the
replicas to be cached. If certain pathologies are discovered (i.e., lost,
unavailable, or zero-sized files on tape), it can update the corresponding
entries to ``Failed'' in a similar manner. Otherwise, they are promoted to
``StageSubmitted''. The agent responsible for monitoring the status of submitted
requests is the \textit{Stage Monitor Agent}. It achieves this by interrogating the
storage middleware to see if the ``StageSubmitted'' files are successfully
replicated on disk cache. In case of success, the \textit{CacheReplicas} and their
corresponding \textit{Tasks} entries are updated to ``Staged''. Various circumstances of
tape or middleware misbehavior can also fail the staging requests. The last one
in the chain is the \textit{Request Finalization Agent}. The \textit{Tasks} which are in their
final states (``Staged'' or ``Failed'') are cleared from the database, and callbacks
are performed to the WMS, which effectively wakes up the corresponding jobs. If
there are no more associated \textit{Tasks} for particular replicas, the respective
\textit{CacheReplicas} entries are also removed. The state machine of the \textit{CacheReplicas}
entity is shown in Fig.~\ref{fig:CacheReplicas-SM}.

@@

\section{Related Work}
\label{sec:RelatedWork}
Model checking belongs to a class of mathematically-based techniques for specification, analysis,
and verification of software and hardware systems. Theorem proving, static checking, run-time verification, and model-based testing
also belong to this category of formal methods. The main premise behind their use is the expectation that
the initial investment in becoming proficient and developing models pays off in the long run, in terms 
of improved reliability and safety of the system. While traditional testing is almost
exclusively used as a verification methodology in many practical industrial cases,
formal methods have started receiving more attention in the last couple of decades. 
Testing is often based on the intuition and experience of the tester, and, as such, does not 
always reveal subtle system errors. Hence, particular efforts are directed at
the practical use of formal methods in the area of critical, concurrent, and distributed 
systems.
The application of such methods in communication protocols~\cite{SPIN_case_study,SlidingWindowProtocol,sessionMgmtMasterThesis}, 
safety-critical software~\cite{Java_PathFinder,CMS_LHC,Mathijssen:2006:VDA:1757571.1757584},
driver design~\cite{Ball04slamand,Linux_driver}, among others, has been a success.
Case studies in the domain of high-performance computing, and in particular concurrent/MPI-based program design,
have surfaced as well~\cite{Siegel04verificationof,Matlin:2002:SPS:645881.672236,Siegel:2007:MCN:1763048.1763053}, mostly 
relying on the the MPI extension of the popular model checker SPIN~\cite{Siegel:2007:VPP:2396095.2396105}.
Ironically enough, the model checking tools themselves do not take advantage of distributed infrastructures, and 
as such do not scale well.

To overcome the fundamental problem of state space explosion, several
variants to explicit model checking have gained popularity. Runtime verification~\cite{citeulike:10198274,Havelund:2008:RVC:1424465.1424470}
aims to increase confidence in the correctness, but does not make claims for absence of defects in a system.
The target program is instrumented manually, or automatically by scanning function calls and variable assignments, 
in order to capture and emit relevant events which are then monitored against a special component during execution.
This component resembles our monitoring process automaton used to cope with state space explosion during model checking 
of DIRAC's WMS system. While runtime verification is useful for partially-capturing incorrect behavior of the system implementation, 
code instrumentation is not always a feasible option, considering the increase of memory consumption. 
Similar to this lightweight form of verification is static program analysis~\cite{Schmidt98programanalysis}. Rather than code instrumentation and execution,
static program analysis aims to determine runtime properties of systems at compilation time.
A special analyzer targeted for the implementation language maintains information on the control flow and function call graphs
of the software. Approximation is used by such tools in practice, in order to improve scalability. The obvious advantage
of this approach is that it works directly on the implementation. However, most static analysis tools aim at discovering 
more general implementation deficiencies, such as uninitialized variables, null pointer references, or out of bound array indices. 
Both runtime verification and static analysis can be seen as additions to
traditional model checking, rather than alternatives. Some progress has been 
made in the direction of combining the approaches~\cite{Java_PathFinder,Linux_driver,Brat01combiningstatic}. For instance, Java PathFinder~\cite{Java_PathFinder}
relies on static analysis to reduce the state space before systematically exploring it by executing the abstracted program. In~\cite{Brat01combiningstatic}
static analysis and model checking also operate in a feedback loop. The drawback of static analyzers is their 
practical availability for languages other than C and Java. To the best of our knowledge, no such tool exists for Python.
However, our approach of using code slicing to abstract the implementation shares similarities.

TODO: trace mining
